---
title: "ENVS 193DS Homework 3"
author: "Stella Symonds"
format: pdf
execute: 
  message: false
  warning: false
---

# Part 1: Set up 

```{r}
#| label: set-up

# reading in packages 
library(tidyverse)
library(janitor)
library(here)
library(readxl)

# reading in data
# kelp data
kelp <- read.csv(
  here("data", "temp-kelp.csv"))
# personal data 
bike_rides <- read_xlsx(
  here("data", "bike-rides.xlsx"))
```

# Part 2: Problems 

## Problem 1 

a. Pearson's correlation and Spearman rank correlation are two tests that determine the strength of the relationship between temperature and giant kelp frond elongation rate. Pearson's correlation requires continuous, normally distributed variables with a linear relationship, while Spearman rank correlation does not require normally distributed variables. 

b. 
```{r}
#| label: visualization # ADD COLOR

# base layer ggplot
ggplot(data = kelp, 
       mapping = aes(x = temp_c, # x axis is temp
                     y = kelp_elong)) + # y axis is kelp growth rate
  # first layer: points
  geom_point(color = "#D6364F", # custom color
             shape = 15) + # make points square
  # rename axes, add units
  labs(x = "Temp (C)",
       y = "Kelp elongation rate (cm/day)") + 
  theme_minimal() # custom theme
```

c. 
### Check assumptions 
```{r}
#| label: check-assumptions

```

### Run the test
```{r}
#| label: run-test

```






